Topology and geometry of the dark matter web: a multistream view

### 2018-03-07 Draft/Chapter4/Paper2017a_Nobmath.tex
initial commit

@@

% \bmath changed to \bf --- check if it's alright. 
The multistream field objectively characterizes the level of non-linearity in the cosmic web. The `number-of-streams' field or $n_{str}({\bmath x})$ is computed from the Lagrangian sub-manifold ${\bmath x}({\bmath q})$, which is a continuous three-dimensional sheet in a six-dimensional  $({\bmath q}, {\bmath x})$ space. In this paper, we utilize the tessellation implementation by \cite{Shandarin2012} to calculate the multistream flow field on the GADGET-2 snapshot at $z=0$. This implementation only requires initial and final coordinates of the dark matter particles.

The $n_{str}(\bmath{x})$ values are mostly odd-numbered since each folding in the Lagrangian sub-manifold results in an increase of $n_{str}$ by 2. Exception to this are only at caustics - which have volume measure zero, then the $n_{str}$ is even-valued number. The particles in $n_{str} = 1$ have not experienced orbit crossings and thus these regions are unambiguously identified as void \citep{Shandarin2012}. Foldings in the Lagrangian sub-manifold generally occur one-by-one. For example, a contour of $n_{str} = 7$ will be within a region of $n_{str} \leq 5$. Hence the multistream field commonly has nesting shells, i.e., $ 3 \supseteq  5 \supseteq  7 \supseteq  9 \supseteq  11 \ldots$. Some of the important features of the multistream field are discussed in Appendix \ref{appendix:nstream}.

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
\centering\includegraphics[width=8.cm]{Chapter4/Source_v2/fig3.pdf} 
\end{minipage}\hfill
\caption{Single-streaming void distribution on diagnostic spheres around FOF-haloes are considered. At radius $r_{10}$, each diagnostic sphere has $n_{str} = 1$ on 10 per cent of its spherical surface. Distribution function of $r_{10}$ (blue) and FOF-radii $r_{vir}$ (red) are shown. Inner plot shows the distribution function of $r_{10}/r_{vir}$. The haloes within the dashed line have at least 10 per cent of their virial-surfaces in contact with $n_{str} = 1$ regions.}
\label{fig:Vfr_all}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{Chapter4/Source_v2/fig4.pdf} 
\end{minipage}\hfill
\caption{Percolation plot in the multistream field and mass density. Two density estimators - CIC and DTFE are shown. Percolation transition (at $f_1/f_{ES} = 0.5$ shown by the horizontal red line) occurs at smaller excursion set volumes for the multistream field, as seen by the dashed lines for both the curves. It is worth stressing that the percolation curves for $n_{str}$ field are bounded by conditions $f_{ES} < 0.1$. }
\label{fig:PercTh}
\end{figure}

@@

\begin{comment}
\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{fig5n.pdf} 
\end{minipage}\hfill
\caption{Percolation threshold in the multistream field with refinement factor of 2 in the simulation with $256^3$ particles. The volume fraction of excursion set and the filling fraction of the largest structure is shown. Percolation transition at $n_{str} = 17$ is shows by the dashed vertical line. }
\label{fig:HaloFilDen}
\end{figure}

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{fig6n.pdf} 
\end{minipage}\hfill
\caption{Percolation threshold in the density field of the simulation with $256^3$ particles calculated from the CIC method. The volume fraction of excursion set and the filling fraction of the largest structure is shown. Percolation at $\rho/ \rho_b = 4.8 $ is shown by the dashed vertical line.}
\label{fig:HaloFilNst}
\end{figure}

@@

In the simulation with $256^3$ particles, percolations in the density field occurs at $\rho_{DTFE}/ \rho_b = 5.16 $ and $\rho_{CIC}/ \rho_b = 5.49 $ for densities calculated with DTFE and CIC respectively. Here $\rho_b = 256^3 / 100^3 M_{\sun} h^{-3} \text{Mpc}^{-3}$, the background density. Notice that these values correspond to the density as calculated by the CIC and DTFE algorithms, and it might be different for other density finding methods. The volume fraction of the excursion set of over-densities at the percolation, $f_{ES}^{(p)} = 2.7$ per cent, is considerably higher than the corresponding $f_{ES}^{(p)}$ value in the multistream field. This implies that the percolation occurs at larger values of filling fraction in mass densities.

@@

\section{Local geometry of the multistream field}
\label{sec:hessian}
\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{Chapter4/Source_v2/fig6.pdf}

\end{minipage}\hfill
\caption{Probability distribution function of the sorted eigenvalues of the Hessian $\mathbfss{H}(-n_{str})$ in the simulation box with $N_p = 128^3$. Top panel: Distribution in the entire simulation box. The multistream field is calculated at refinement factor $l_l/l_d= 2$ and smoothing scale of equal to $l_d$. All the three eigenvalue data fields have a highest number of points where their value is 0. Bottom panel: Hessian eigenvalues for the non-void region ($n_{str} > 1$) is shown. Total number of eigenvalue triplets are less than $10$ per cent of that of the full simulation box. Eigenvalues close to zero in non-void regions are notably fewer than in the entire simulation box.}
\label{fig:lambdasPDF}
\end{figure}

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{Chapter4/Source_v2/fig7.pdf}

\end{minipage}\hfill
\caption{Comparison between small eigenvalues of the multistream Hessian $\mathbfss{H}(-n_{str})$. Percentage of eigenvalues with absolute values less than a cut-off, $\lambda_{th}$ are shown for full simulation box (dashed lines) and the multistream web structure (solid lines). The multistream web has fewer eigenvalues below $\lambda_{th} = 0.1$. The void seems to have most of the small eigenvalues. } 
\label{fig:lambdasSmall}
\end{figure}

@@

Local second order variation in a scalar field $f$ like the multistream field can by found using the Hessian matrix $\mathbfss{H}(f)$. An element of the Hessian matrix is given in \autoref{eq:Hess}, where $i$ and $j$ can be any of $x$, $y$ or $z$ directions.

\begin{equation}
\label{eq:Hess}
\mathbfss{H}_{ij}(f) = \frac{\partial^2 f}{\partial x_i \partial x_j} 
\end{equation}

In our analysis, we have chosen $f = -n_{str}(\bmath{x})$ for understanding local variations of the multistream field. The resulting Hessians at each point on the configuration space are always symmetric matrices, as illustrated in Appendix \ref{appendix:Eigen}. The eigenvalues of these Hessian matrices are always real, and depending on if their values are positive or negative, one may infer local geometrical features in the multistream field. 
\begin{comment}
We utilise a framework based on Hessian eigenvalues in smoothed multistream field for the detection of haloes.   
\end{comment}

Within the void, there is no variation in the multistream values. Hessians $\mathbfss{H}(-n_{str})$ are zero matrices in large volume fraction of the simulation box (around 90 per cent in both the simulations) due to the constant value of $n_{str} = 1$ in this percolating void. Eigenvalues of these Hessian matrices, sorted as $ \lambda_1 \geq \lambda_2 \geq \lambda_3 $ are close to 0 at a large number of regions as shown in the top panel of \autoref{fig:lambdasPDF}. In the simulation with $128^3$ particles, the median values of each eigenvalue are $0.09$, $-3\times 10^{-10}$ and $-0.11$ for  $\lambda_1$, $\lambda_2$ and $\lambda_3$ respectively. By selecting just the non-void region by $n_{str} > 1$, notably fewer number of eigenvalues have small absolute values. The median values of each of the eigenvalues in the non-void regions are $4.01$, $0.48$, and $-0.85$ respectively for $\lambda_1$, $\lambda_2$ and  $\lambda_3$. Bottom panel in \autoref{fig:lambdasPDF} shows a significant change in the probability distribution of Hessian eigenvalues around 0, the distribution pattern at the tails are mostly identical to the distribution pattern in the entire simulation box.

@@

Hessian eigenvalues in multistream fields differ from that in density, gravitational potential or velocity shear tensor. Constant scalar value of $n_{str}$ facilitates the Hessian $\mathbfss{H}(-n_{str})$ matrices to be presumptively close to zero. On the other hand, in density field manifests in a range of low values in the voids, resulting in non-zero Hessian matrices. Eigenvalues of velocity shear tensor do not peak at zero either \cite{Libeskind2013}. For the deformation tensor, morphological characterization of the cosmic web using Zel'dovich formalism shows that each eigenvalue must be negative in voids.

\begin{figure*}
\begin{minipage}[t]{0.99\linewidth}
 \centering\includegraphics[height=14.cm]{Chapter4/Source_v2/fig8.pdf} 
\end{minipage}\hfill
\caption{Eigenvalues of the Hessian matrix $\mathbfss{H}(-n_{str})$ in a slice of 50 $h^{-1}$ Mpc $\times$ 50 $h^{-1}$ Mpc slice of the simulation box of $128^3$ particles. Variation in the eigenvalues in the multistreaming web structure is shown. The largest eigenvalue $\lambda_1$ (top left panel) has positive values throughout the structure. The smallest eigenvalue $\lambda_3$ (bottom left) has negative values surrounding positive definite regions of the $n_{str}$ field. Corresponding multistream field is shown in the bottom right panel for single, three and more than five streams.}
\label{fig:evals123}
\end{figure*}

@@

The eigenvalues of $\mathbfss{H}(-n_{str})$ span a large range of values in our cosmological simulation. The largest eigenvalue of the triplets, $\lambda_1$ having large positive values throughout the multistream web structure ( see \autoref{fig:evals123} ). Absolute values $|\lambda_1|$, $|\lambda_2|$ and $|\lambda_3|$ peak around the neighbourhood of intersections of filaments. These junctions are usually high streaming regions due to shell crossing from multiple directions. \cite{Ramachandra2015} observed that these regions with intersecting filaments are in the vicinity of large FOF haloes.

\begin{comment}
\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{fig5.pdf}

@@

If the Hessian matrices are positive definite in a region, i.e., if all the eigenvalues are strictly positive, then the interior of this convex region has at-most one minimum. For our choice of $-n_{str}(\bmath{x})$ as the domain of Hessian, this means that the convex neighbourhoods around local maxima of the multistream field are isolated by the positive definite Hessian matrices. Closed surface contours at high streaming or the most non-linear regions are selected. These regions my indeed be the regions of dark matter haloes.

@@

Hessian eigenvalues are generally defined on continuous functions. Although our domain of the Hessian is an inhererntly integer-valued field, it describes the multistream structure at the level of diagnostic grid. Hence it may be considered to be numerically equivalent to a continuous function where the numerical approxiamtion of differentation is a valid operation. This can be verified mathematically by finding that Hessian $\mathbfss{H}(-n_{str})$ is symmetric (Appendix \ref{appendix:Eigen} shows the numerical approximation of the Hessian matrix term for generic unfiltered multistream field.)

@@

\begin{comment}
\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{fig9.pdf} 
\end{minipage}\hfill
\caption{The distribution of $\lambda_3$ in the simulation box of $128^3$ particles and multistream field of refinement factor $l_l/l_d = 2$. Three smoothing scales are shown. }
\label{fig:Eval3Smooth}
\end{figure}
\end{comment}

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{Chapter4/Source_v2/fig10.pdf}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{Chapter4/Source_v2/fig11.pdf}

@@

However, local geometry analysis of the multistream flow field varies considerably on the resolution of the analysis grid. For our Hessian $\mathbfss{H}(-n_{str})$, the regions with $\lambda_1 \geq \lambda_2 \geq \lambda_3 > 0$ in non-void regions occupy 1.8 per cent of the entire box in native resolution of diagnostic grid, as shown in \autoref{tab:VolfrRefLambda}. This fraction reduces to 1.3 per cent at diagnostic grid of $512^3$ resolution. Variations with refinement factors are seen in other eigenvalue conditions in the non-void too: volume fraction of $\lambda_1 > 0 > \lambda_2 \geq \lambda_3 $ regions increases from 1.7 per cent at refinement factor of 1 to 3 per cent at refinement factor of 8. Volume fraction of $\lambda_1 \geq \lambda_2 > 0 > \lambda_3 $ regions decreases from 5.6 to 4.6 per cent with the increase of refinement from 1 to 8.

@@

\begin{table}
  \caption{Volume fraction of criteria based on $n_{str}$ and $\lambda$s of $\mathbfss{H}(-n_{str})$ calculated at various resolutions. We chose a smaller slice of $ 50 h^{-1} \text{Mpc} \times 50 h^{-1} \text{Mpc} \times 50 h^{-1} \text{Mpc} $ i.e., half the volume of the original GADGET simulation box. The refinement factors are the multiplication factors of 1, 2, 4 and 8 times of the native resolution ($64^3$) of the simulation grid along each axis. Eigenvalues of the Hessian of the field are local geometric parameters. The void is globally defined as $n_{str} = 1$ and the multistream web structure as $n_{str} > 1$.}

@@

We intend to identify haloes in the $n_{str}(\bmath{x})$ field instead of using just the position coordinate data. While the eigenvalue analysis itself is done at a chosen time, the multistream field inherently has data from six-dimensional Lagrangian space $(\bmath{q}, \bmath{x})$ that contains the full dynamical information,similar to the phase-space sheet albeit in a different form. Dynamical history that is embedded in the multistream field is crucial in understanding the strength of gravitational binding of the particles. 
Moreover, as explained in Section \ref{sec:voids}, a physically motivated distinction between void and gravitationally collapsed regions is a unique feature of multistream analysis. Thus the haloes detected from local maxima of the $n_{str}$ field can be ensured to be away from the single streaming voids. Methods based on linking-length or density fields may not able to ensure that all the particles in haloes are away from voids, as shown for FOF method in Section \ref{sec:voidHaloes}.

@@

\item Second order variations of the smoothed $-n_{str}(\bmath{x})$ is computed at each point in the field. This gives symmetric Hessian matrices for this field whose eigenvalues are real. Ordered eigenvalues of the Hessian, $ \lambda_1 \geq  \lambda_2 \geq \lambda_3$ are calculated. The $\lambda_3$ field is shown in the top right panel of \autoref{fig:labelsfilter}.

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{fig11.pdf} 
\end{minipage}\hfill
\caption{PDF of highest $\lambda_1$, $\lambda_2$ and $\lambda_3$ values in each of 4492 haloes detected by out algorithm. The peaks of the PDF are in the range 1-10. Shaded regions represent 1$\sigma$ error. }
\label{fig:maxL3}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{fig12.pdf} 
\end{minipage}\hfill
\caption{Maximum, minimum and median of $n_{str}$ in each of 4492 halo candidates. The closed contours of haloes encompass a wide range of multistream values. None of the haloes are in contact with the void region, since lowest value of min$(n_{str})$ is 3. Shaded regions are the 1$\sigma$ absolute errors in the number of $\lambda_3$-haloes. }
\label{fig:minmaxstr}
\end{figure}

@@

However, the $\lambda_3$-halo boundary is different from any constant multistream contour. That is, from the function $n_{str}(\bmath{x})$, convex regions in the four-dimensional function space $(-n_{str}, x, y, z)$ are projected onto three-dimensional co-ordinate space using eigenvalues. This is different from projecting `iso-multistream' slice onto three-dimensional co-ordinate space. Appendix \ref{appendix:Eigen} illustrates the difference in the two approaches for a one-dimensional function.

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering
 \includegraphics[width=8.cm]{fig14}
\end{minipage}\hfill
\caption{Mass fractions (in per cent) of haloes detected by the three finders. The particles concurrently detected as belonging to haloes are shown in the intersections. }
\label{fig:HaloFinderMF}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{fig15.pdf} 
\end{minipage}\hfill
\caption{Halo Mass functions from AHF-, FOF- and $\lambda_3$-haloes. The AHF-haloes are fewer than FOF- or $\lambda_3$-haloes. The number of haloes above a mass threshold are binned and taken along vertical axis, normalized to simulation box volume. Error of 1$\sigma$ is shown in shaded region. }
\label{fig:hmf}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering
 \includegraphics[width=8.cm]{fig16.pdf}
\end{minipage}\hfill
\caption{A large halo that was detected by AHF (red), FOF (blue) and our geometric analysis in the multistream field (green). Halo boundary differs for each halo finder method. AHF detects particles within a sphere of virial radius. FOF-halo is irregularly shaped. $\lambda_3$-halo particles are in a non-spherical, yet compact structure. }
\label{fig:FinderCompareAll}
\end{figure}

@@

\begin{table}
  \caption{Number of $\lambda_3$-haloes identified at smoothing of $n_{str}(\bmath{x})$ at different scales.}
\begin{tabular}{|l|r|r|r|}
\hline
$N_p$  &  $0.20h^{-1} \text{Mpc}$ & $0.39 h^{-1} \text{Mpc}$ & $0.78 h^{-1} \text{Mpc}$ \\ \hline
$128^3$   & $5181$  &  $4492$ & $2923$ \\ \hline
$256^3$   & $27929$  & $18221$ & $7897$ \\ \hline

>>>>

% \mathbf -> \mathbf
% \mathbfss{H} -> {\bf H}
% \sun -> \odot

@@

% \bf changed to \bf --- check if it's alright. 
The multistream field objectively characterizes the level of non-linearity in the cosmic web. The `number-of-streams' field or $n_{str}({\bf x})$ is computed from the Lagrangian sub-manifold ${\bf x}({\bf q})$, which is a continuous three-dimensional sheet in a six-dimensional  $({\bf q}, {\bf x})$ space. In this paper, we utilize the tessellation implementation by \cite{Shandarin2012} to calculate the multistream flow field on the GADGET-2 snapshot at $z=0$. This implementation only requires initial and final coordinates of the dark matter particles.

The $n_{str}(\bf{x})$ values are mostly odd-numbered since each folding in the Lagrangian sub-manifold results in an increase of $n_{str}$ by 2. Exception to this are only at caustics - which have volume measure zero, then the $n_{str}$ is even-valued number. The particles in $n_{str} = 1$ have not experienced orbit crossings and thus these regions are unambiguously identified as void \citep{Shandarin2012}. Foldings in the Lagrangian sub-manifold generally occur one-by-one. For example, a contour of $n_{str} = 7$ will be within a region of $n_{str} \leq 5$. Hence the multistream field commonly has nesting shells, i.e., $ 3 \supseteq  5 \supseteq  7 \supseteq  9 \supseteq  11 \ldots$. Some of the important features of the multistream field are discussed in Appendix \ref{appendix:nstream}.

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
\centering\includegraphics[width=10.cm]{Chapter4/Source_v2/fig3.pdf} 
\end{minipage}\hfill
\caption{Single-streaming void distribution on diagnostic spheres around FOF-haloes are considered. At radius $r_{10}$, each diagnostic sphere has $n_{str} = 1$ on 10 per cent of its spherical surface. Distribution function of $r_{10}$ (blue) and FOF-radii $r_{vir}$ (red) are shown. Inner plot shows the distribution function of $r_{10}/r_{vir}$. The haloes within the dashed line have at least 10 per cent of their virial-surfaces in contact with $n_{str} = 1$ regions.}
\label{fig:Vfr_all}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{Chapter4/Source_v2/fig4.pdf} 
\end{minipage}\hfill
\caption{Percolation plot in the multistream field and mass density. Two density estimators - CIC and DTFE are shown. Percolation transition (at $f_1/f_{ES} = 0.5$ shown by the horizontal red line) occurs at smaller excursion set volumes for the multistream field, as seen by the dashed lines for both the curves. It is worth stressing that the percolation curves for $n_{str}$ field are bounded by conditions $f_{ES} < 0.1$. }
\label{fig:PercTh}
\end{figure}

@@

\begin{comment}
\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{fig5n.pdf} 
\end{minipage}\hfill
\caption{Percolation threshold in the multistream field with refinement factor of 2 in the simulation with $256^3$ particles. The volume fraction of excursion set and the filling fraction of the largest structure is shown. Percolation transition at $n_{str} = 17$ is shows by the dashed vertical line. }
\label{fig:HaloFilDen}
\end{figure}

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{fig6n.pdf} 
\end{minipage}\hfill
\caption{Percolation threshold in the density field of the simulation with $256^3$ particles calculated from the CIC method. The volume fraction of excursion set and the filling fraction of the largest structure is shown. Percolation at $\rho/ \rho_b = 4.8 $ is shown by the dashed vertical line.}
\label{fig:HaloFilNst}
\end{figure}

@@

In the simulation with $256^3$ particles, percolations in the density field occurs at $\rho_{DTFE}/ \rho_b = 5.16 $ and $\rho_{CIC}/ \rho_b = 5.49 $ for densities calculated with DTFE and CIC respectively. Here $\rho_b = 256^3 / 100^3 M_{\odot} h^{-3} \text{Mpc}^{-3}$, the background density. Notice that these values correspond to the density as calculated by the CIC and DTFE algorithms, and it might be different for other density finding methods. The volume fraction of the excursion set of over-densities at the percolation, $f_{ES}^{(p)} = 2.7$ per cent, is considerably higher than the corresponding $f_{ES}^{(p)}$ value in the multistream field. This implies that the percolation occurs at larger values of filling fraction in mass densities.

@@

\section{Local geometry of the multistream field}
\label{sec:hessian}
\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{Chapter4/Source_v2/fig6.pdf}

\end{minipage}\hfill
\caption{Probability distribution function of the sorted eigenvalues of the Hessian ${\bf H}(-n_{str})$ in the simulation box with $N_p = 128^3$. Top panel: Distribution in the entire simulation box. The multistream field is calculated at refinement factor $l_l/l_d= 2$ and smoothing scale of equal to $l_d$. All the three eigenvalue data fields have a highest number of points where their value is 0. Bottom panel: Hessian eigenvalues for the non-void region ($n_{str} > 1$) is shown. Total number of eigenvalue triplets are less than $10$ per cent of that of the full simulation box. Eigenvalues close to zero in non-void regions are notably fewer than in the entire simulation box.}
\label{fig:lambdasPDF}
\end{figure}

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{Chapter4/Source_v2/fig7.pdf}

\end{minipage}\hfill
\caption{Comparison between small eigenvalues of the multistream Hessian ${\bf H}(-n_{str})$. Percentage of eigenvalues with absolute values less than a cut-off, $\lambda_{th}$ are shown for full simulation box (dashed lines) and the multistream web structure (solid lines). The multistream web has fewer eigenvalues below $\lambda_{th} = 0.1$. The void seems to have most of the small eigenvalues. } 
\label{fig:lambdasSmall}
\end{figure}

@@

Local second order variation in a scalar field $f$ like the multistream field can by found using the Hessian matrix ${\bf H}(f)$. An element of the Hessian matrix is given in \autoref{eq:Hess}, where $i$ and $j$ can be any of $x$, $y$ or $z$ directions.

\begin{equation}
\label{eq:Hess}
 {\bf H}_{ij}(f) = \frac{\partial^2 f}{\partial x_i \partial x_j} 
\end{equation}

In our analysis, we have chosen $f = -n_{str}(\bf{x})$ for understanding local variations of the multistream field. The resulting Hessians at each point on the configuration space are always symmetric matrices, as illustrated in Appendix \ref{appendix:Eigen}. The eigenvalues of these Hessian matrices are always real, and depending on if their values are positive or negative, one may infer local geometrical features in the multistream field. 
\begin{comment}
We utilise a framework based on Hessian eigenvalues in smoothed multistream field for the detection of haloes.   
\end{comment}

Within the void, there is no variation in the multistream values. Hessians ${\bf H}(-n_{str})$ are zero matrices in large volume fraction of the simulation box (around 90 per cent in both the simulations) due to the constant value of $n_{str} = 1$ in this percolating void. Eigenvalues of these Hessian matrices, sorted as $ \lambda_1 \geq \lambda_2 \geq \lambda_3 $ are close to 0 at a large number of regions as shown in the top panel of \autoref{fig:lambdasPDF}. In the simulation with $128^3$ particles, the median values of each eigenvalue are $0.09$, $-3\times 10^{-10}$ and $-0.11$ for  $\lambda_1$, $\lambda_2$ and $\lambda_3$ respectively. By selecting just the non-void region by $n_{str} > 1$, notably fewer number of eigenvalues have small absolute values. The median values of each of the eigenvalues in the non-void regions are $4.01$, $0.48$, and $-0.85$ respectively for $\lambda_1$, $\lambda_2$ and  $\lambda_3$. Bottom panel in \autoref{fig:lambdasPDF} shows a significant change in the probability distribution of Hessian eigenvalues around 0, the distribution pattern at the tails are mostly identical to the distribution pattern in the entire simulation box.

@@

Hessian eigenvalues in multistream fields differ from that in density, gravitational potential or velocity shear tensor. Constant scalar value of $n_{str}$ facilitates the Hessian ${\bf H}(-n_{str})$ matrices to be presumptively close to zero. On the other hand, in density field manifests in a range of low values in the voids, resulting in non-zero Hessian matrices. Eigenvalues of velocity shear tensor do not peak at zero either \cite{Libeskind2013}. For the deformation tensor, morphological characterization of the cosmic web using Zel'dovich formalism shows that each eigenvalue must be negative in voids.

\begin{figure*}
\begin{minipage}[t]{0.99\linewidth}
 \centering\includegraphics[height=14.cm]{Chapter4/Source_v2/fig8.pdf} 
\end{minipage}\hfill
\caption{Eigenvalues of the Hessian matrix ${\bf H}(-n_{str})$ in a slice of 50 $h^{-1}$ Mpc $\times$ 50 $h^{-1}$ Mpc slice of the simulation box of $128^3$ particles. Variation in the eigenvalues in the multistreaming web structure is shown. The largest eigenvalue $\lambda_1$ (top left panel) has positive values throughout the structure. The smallest eigenvalue $\lambda_3$ (bottom left) has negative values surrounding positive definite regions of the $n_{str}$ field. Corresponding multistream field is shown in the bottom right panel for single, three and more than five streams.}
\label{fig:evals123}
\end{figure*}

@@

The eigenvalues of ${\bf H}(-n_{str})$ span a large range of values in our cosmological simulation. The largest eigenvalue of the triplets, $\lambda_1$ having large positive values throughout the multistream web structure ( see \autoref{fig:evals123} ). Absolute values $|\lambda_1|$, $|\lambda_2|$ and $|\lambda_3|$ peak around the neighbourhood of intersections of filaments. These junctions are usually high streaming regions due to shell crossing from multiple directions. \cite{Ramachandra2015} observed that these regions with intersecting filaments are in the vicinity of large FOF haloes.

\begin{comment}
\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{fig5.pdf}

@@

If the Hessian matrices are positive definite in a region, i.e., if all the eigenvalues are strictly positive, then the interior of this convex region has at-most one minimum. For our choice of $-n_{str}(\bf{x})$ as the domain of Hessian, this means that the convex neighbourhoods around local maxima of the multistream field are isolated by the positive definite Hessian matrices. Closed surface contours at high streaming or the most non-linear regions are selected. These regions my indeed be the regions of dark matter haloes.

@@

Hessian eigenvalues are generally defined on continuous functions. Although our domain of the Hessian is an inhererntly integer-valued field, it describes the multistream structure at the level of diagnostic grid. Hence it may be considered to be numerically equivalent to a continuous function where the numerical approxiamtion of differentation is a valid operation. This can be verified mathematically by finding that Hessian ${\bf H}(-n_{str})$ is symmetric (Appendix \ref{appendix:Eigen} shows the numerical approximation of the Hessian matrix term for generic unfiltered multistream field.)

@@

\begin{comment}
\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=10.cm]{fig9.pdf} 
\end{minipage}\hfill
\caption{The distribution of $\lambda_3$ in the simulation box of $128^3$ particles and multistream field of refinement factor $l_l/l_d = 2$. Three smoothing scales are shown. }
\label{fig:Eval3Smooth}
\end{figure}
\end{comment}

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{Chapter4/Source_v2/fig10.pdf}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{Chapter4/Source_v2/fig11.pdf}

@@

However, local geometry analysis of the multistream flow field varies considerably on the resolution of the analysis grid. For our Hessian ${\bf H}(-n_{str})$, the regions with $\lambda_1 \geq \lambda_2 \geq \lambda_3 > 0$ in non-void regions occupy 1.8 per cent of the entire box in native resolution of diagnostic grid, as shown in \autoref{tab:VolfrRefLambda}. This fraction reduces to 1.3 per cent at diagnostic grid of $512^3$ resolution. Variations with refinement factors are seen in other eigenvalue conditions in the non-void too: volume fraction of $\lambda_1 > 0 > \lambda_2 \geq \lambda_3 $ regions increases from 1.7 per cent at refinement factor of 1 to 3 per cent at refinement factor of 8. Volume fraction of $\lambda_1 \geq \lambda_2 > 0 > \lambda_3 $ regions decreases from 5.6 to 4.6 per cent with the increase of refinement from 1 to 8.

@@

\begin{table}
  \caption{Volume fraction of criteria based on $n_{str}$ and $\lambda$s of ${\bf H}(-n_{str})$ calculated at various resolutions. We chose a smaller slice of $ 50 h^{-1} \text{Mpc} \times 50 h^{-1} \text{Mpc} \times 50 h^{-1} \text{Mpc} $ i.e., half the volume of the original GADGET simulation box. The refinement factors are the multiplication factors of 1, 2, 4 and 8 times of the native resolution ($64^3$) of the simulation grid along each axis. Eigenvalues of the Hessian of the field are local geometric parameters. The void is globally defined as $n_{str} = 1$ and the multistream web structure as $n_{str} > 1$.}

@@

We intend to identify haloes in the $n_{str}(\bf{x})$ field instead of using just the position coordinate data. While the eigenvalue analysis itself is done at a chosen time, the multistream field inherently has data from six-dimensional Lagrangian space $(\bf{q}, \bf{x})$ that contains the full dynamical information,similar to the phase-space sheet albeit in a different form. Dynamical history that is embedded in the multistream field is crucial in understanding the strength of gravitational binding of the particles. 
Moreover, as explained in Section \ref{sec:voids}, a physically motivated distinction between void and gravitationally collapsed regions is a unique feature of multistream analysis. Thus the haloes detected from local maxima of the $n_{str}$ field can be ensured to be away from the single streaming voids. Methods based on linking-length or density fields may not able to ensure that all the particles in haloes are away from voids, as shown for FOF method in Section \ref{sec:voidHaloes}.

@@

\item Second order variations of the smoothed $-n_{str}(\bf{x})$ is computed at each point in the field. This gives symmetric Hessian matrices for this field whose eigenvalues are real. Ordered eigenvalues of the Hessian, $ \lambda_1 \geq  \lambda_2 \geq \lambda_3$ are calculated. The $\lambda_3$ field is shown in the top right panel of \autoref{fig:labelsfilter}.

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=10.cm]{fig11.pdf} 
\end{minipage}\hfill
\caption{PDF of highest $\lambda_1$, $\lambda_2$ and $\lambda_3$ values in each of 4492 haloes detected by out algorithm. The peaks of the PDF are in the range 1-10. Shaded regions represent 1$\sigma$ error. }
\label{fig:maxL3}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=10.cm]{fig12.pdf} 
\end{minipage}\hfill
\caption{Maximum, minimum and median of $n_{str}$ in each of 4492 halo candidates. The closed contours of haloes encompass a wide range of multistream values. None of the haloes are in contact with the void region, since lowest value of min$(n_{str})$ is 3. Shaded regions are the 1$\sigma$ absolute errors in the number of $\lambda_3$-haloes. }
\label{fig:minmaxstr}
\end{figure}

@@

However, the $\lambda_3$-halo boundary is different from any constant multistream contour. That is, from the function $n_{str}(\bf{x})$, convex regions in the four-dimensional function space $(-n_{str}, x, y, z)$ are projected onto three-dimensional co-ordinate space using eigenvalues. This is different from projecting `iso-multistream' slice onto three-dimensional co-ordinate space. Appendix \ref{appendix:Eigen} illustrates the difference in the two approaches for a one-dimensional function.

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering
 \includegraphics[width=10.cm]{fig14}
\end{minipage}\hfill
\caption{Mass fractions (in per cent) of haloes detected by the three finders. The particles concurrently detected as belonging to haloes are shown in the intersections. }
\label{fig:HaloFinderMF}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=10.cm]{fig15.pdf} 
\end{minipage}\hfill
\caption{Halo Mass functions from AHF-, FOF- and $\lambda_3$-haloes. The AHF-haloes are fewer than FOF- or $\lambda_3$-haloes. The number of haloes above a mass threshold are binned and taken along vertical axis, normalized to simulation box volume. Error of 1$\sigma$ is shown in shaded region. }
\label{fig:hmf}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering
 \includegraphics[width=10.cm]{fig16.pdf}
\end{minipage}\hfill
\caption{A large halo that was detected by AHF (red), FOF (blue) and our geometric analysis in the multistream field (green). Halo boundary differs for each halo finder method. AHF detects particles within a sphere of virial radius. FOF-halo is irregularly shaped. $\lambda_3$-halo particles are in a non-spherical, yet compact structure. }
\label{fig:FinderCompareAll}
\end{figure}

@@

\begin{table}
  \caption{Number of $\lambda_3$-haloes identified at smoothing of $n_{str}(\bf{x})$ at different scales.}
\begin{tabular}{|l|r|r|r|}
\hline
$N_p$  &  $0.20h^{-1} \text{Mpc}$ & $0.39 h^{-1} \text{Mpc}$ & $0.78 h^{-1} \text{Mpc}$ \\ \hline
$128^3$   & $5181$  &  $4492$ & $2923$ \\ \hline
$256^3$   & $27929$  & $18221$ & $7897$ \\ \hline
